---
title: "Buffered Random Sampling"
author: "Johan Ejstrud"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Buffered Random Sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package implements the buffered random sampling algorithm described in a [paper by Kingsley, M., Kanneworff, P., & Carlsson, D. (2004)](https://academic.oup.com/icesjms/article/61/1/12/699744).

# Getting started
The main function of the package, `buffered_random_sampling`, requires two data sets: one with stratums and one with elements (points) within the stratums.

The package include two example data sets, `element` and `stratum`, that shows how they should look. (See the exact details by running `?element` and `?stratum` in the console.)

```{r setup}
library(buffersam)
set.seed(92)

stratum
head(element)
```

To create an allocation run the function on two data sets:
```{r}
allocation <- buffered_random_sampling(element, stratum)
head(allocation)
```

# Visualise allocation
It's possible to visualise the allocation directly in R, by using `visualise_allocation`.

```{r eval=FALSE}
visualise_allocation(allocation)
```

To see the algorithm as it runs, use `visualise=TRUE`.

```{r eval=FALSE}
buffered_random_sampling(element, stratum, visualise="stratum")
```

The algorithm can stop with different frequencies. See `?buffered_random_sampling` for all options.

# Algorithm in words
This section contains a brief descriptions of the buffered random algorithm in words. Please refer to the paper for exact details.

1. Create a list of all stratums, and a list of all elements.
1. Calculate the sampling density for each stratum by dividing the required number of stations with the stratum's area.
1. Pick the stratum with the lowest sampling density as the _current stratum_.
1. Set buffering distance to sqrt((2A)/(nπ)), where A is stratum area and n is the required number of stations for the current startum.
1. Pick an element at random that is not closer than the buffering distance to any other selected elements.
    1. If the element is within the current stratum, mark it as  _selected_.
    1. If the element is not within the current startum, mark it as _temporary selected_.     
1. If the required number of stations has been reached:
    1. Remove the current stratum from the stratum list.
    1. Move all the temporary selected elements back to the element list.
    1. Go to 3.
1. If the element list is empty:
    1. Move all the temporary selected elements back to the element list.
    1. Move all the selected elements from this stratum back to the element list.
    1. Reduce the buffering distance by 10 %.
    1. Go to 5.
1. Go to 5.


## Stratums and Elements
The _survey area_ is divided into a number of _stratums_ which contains a number of _elements_. Each element is a fixed point with a surrounding area. In the West Greenland shrimp survey each element is the center point of an approximately 2-n.mi. square.

The algorithm doesn't impose any requirements on how the elements are laid out, or how many of them are avaiable. However, an element list is needed; the algorithm cannot select arbitrary points within each stratum.
