---
title: "Buffered Random Sampling"
author: "Johan Ejstrud"
output:
  rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{Buffered Random Sampling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This package implements the buffered random sampling algorithm described in a [paper by Kingsley, M., Kanneworff, P., & Carlsson, D. (2004)](https://academic.oup.com/icesjms/article/61/1/12/699744).

# Getting started
The main function of the package, `buffered_random_sampling`, requires two data sets: one with stratums and one with elements (points) within the stratums.

The package include two example data sets, `element` and `stratum`, that shows how they should look. (See the exact details by running `?element` and `?stratum` in the console.)

```{r setup}
library(buffersam)
set.seed(92)

stratum
head(element)
```

To create an allocation run the function on two data sets:
```{r}
allocation <- buffered_random_sampling(element, stratum)
allocation
```

# Get feedback while algorithm is running
It's possible to get feedback while the algorithm is running, to know far it has come, or to better understand how the algorithm works. This feedback can be provided in text of visuals.

## Visual feedback
It's possible to visualise the allocation by using `visualise=TRUE`.

```{r}
allocation <- buffered_random_sampling(element, stratum, visualise=TRUE)
```


## Text feedback
Use `verbose=TRUE` to print the status in the terminal as the algorithm is running.

```{r}
allocation <- buffered_random_sampling(element, stratum, verbose=TRUE)
```

## Pause algorithm
It's possible to pause the execution of the algorithm while it's running to allow time for inspecting the progress. This only really makes sense if used together with one of the other feedback parameters.

```{r eval=FALSE}
allocation <- buffered_random_sampling(element, stratum, pause=TRUE)
```
## Level of detail
`detail` controls how often the algorithm should give feedback, visual, text, or
pause. Level 1 for least detail, level 4 for most.
```{r}
allocation <- buffered_random_sampling(element, stratum, verbose=TRUE, detail=2)
```

```{r}
allocation <- buffered_random_sampling(element, stratum, verbose=TRUE, detail=3)
```

# Preselect elements
It's possible that some elements need to be selected without regards to buffered random sampling. These elements should be pre-selected, as if the algorithm had already selected them. This could be elements from previous years that need to be repeated.

```{r}
preselect_element <- head(element, 2)
preselect_element
```
```{r eval=FALSE}
allocation <- buffered_random_sampling(element, stratum, preselect_element = preselect_element)
```

# Algorithm in words
This section contains a brief descriptions of the buffered random algorithm in words. Please refer to the paper for exact details.

1. Create a list of all stratums, and a list of all elements.
1. Calculate the sampling density for each stratum by dividing the required number of stations with the stratum's area.
1. Pick the stratum with the lowest sampling density as the _current stratum_.
1. Set buffering distance to sqrt((2A)/(nπ)), where A is stratum area and n is the required number of stations for the current startum.
1. Pick an element at random that is not closer than the buffering distance to any other selected elements.
    1. If the element is within the current stratum, mark it as  _selected_.
    1. If the element is not within the current startum, mark it as _temporary selected_.     
1. If the required number of stations has been reached:
    1. Remove the current stratum from the stratum list.
    1. Move all the temporary selected elements back to the element list.
    1. Go to 3.
1. If the element list is empty:
    1. Move all the temporary selected elements back to the element list.
    1. Move all the selected elements from this stratum back to the element list.
    1. Reduce the buffering distance by 10 %.
    1. Go to 5.
1. Go to 5.


## Stratums and Elements
The _survey area_ is divided into a number of _stratums_ which contains a number of _elements_. Each element is a fixed point with a surrounding area. In the West Greenland shrimp survey each element is the center point of an approximately 2 nautical miles square.

The algorithm doesn't impose any requirements on how the elements are laid out, or how many of them are available. However, an element list is needed; the algorithm cannot select arbitrary points within each stratum.

## Elements belonging to multiple strata
Elements can be placed so its surrounding area is inside two or more strata. This can be represented in the `elements` data frame by having multiple rows with the same `elementId` butna different values for `stratum`. 

Before the algorithm runs, these elements are assigned to _one_ stratum at random, by removing all but one of the duplicated `elementId`'s from the data set. If the `elements` data frame contains the column `preselect_probability` the sampling will be done proportionally to the values in this column. A possible value for this variable is the area the element occupy in the given stratum.

This is done to avoid increasing the likelihood of these elements being selected. Pre-selected elements are always kept.

This peculiarity is not described in the original paper. 
